### 2020-09-28 Latex/Manuscript.pandoc.tex
md conversion to latex #24
- added md2latex.cmd
- Manuscript.pandoc.tex is auto-generated and must be refined

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\date{}

\begin{document}

\hypertarget{document-header}{%
\section{Document Header}\label{document-header}}

\hypertarget{title-proposals}{%
\subsection{Title proposals}\label{title-proposals}}

\begin{itemize}
\tightlist
\item
  Object-Oriented Internet - reactive visualization of asynchronous data
  using AZURE
\item
  Object-Oriented Internet - Azure interoperability
\end{itemize}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

\hypertarget{key-words}{%
\subsection{Key words}\label{key-words}}

Azure, Cloud Computing, Object-Oriented Internet, OPC Unified
Architecture, Machine to Machine Communication, Internet of Things, OPC
UA, RxNetworking

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{itemize}
\tightlist
\item
  \textbf{Subject} - A basic matter of thought, discussion,
  investigation, development, etc. Describe the problem and the
  motivation for undertaking the effort to solve the problem.
\item
  \textbf{Goal} What we are going to achieve - the result or achievement
  toward which effort is directed.
\item
  \textbf{Scope} - What we must do to prove the goal have been achieved.
  Extent or range of development, view, outlook, application, operation,
  effectiveness, etc.
\item
  \textbf{Related work} - Any information about available reusable
  deliverables related to this work.
\end{itemize}

\hypertarget{azure-main-technology-features}{%
\section{Azure Main Technology
Features}\label{azure-main-technology-features}}

\begin{itemize}
\tightlist
\item
  \textbf{Selection of the service}
\item
  \textbf{Metadata} - must be discussed in context of the design/run
  time stages.

  \begin{itemize}
  \tightlist
  \item
    \textbf{Device Template (DT)}
  \item
    \textbf{Device Capability Model}
  \item
    \textbf{Interface}
  \item
    \textbf{Digital Twin Definition Language (DTDL)}
  \end{itemize}
\item
  \textbf{Simple, complex and structural data processing}
\item
  \textbf{Connectivity}
\item
  \textbf{How to implement} All about available libraries and tools
\end{itemize}

\hypertarget{ooi-main-technology-features}{%
\section{OOI Main Technology
Features}\label{ooi-main-technology-features}}

\begin{itemize}
\tightlist
\item
  Machinie To Machine communication based on the semantic data
\item
  OOI PubSub Implementation Architecture
\item
  Simple, complex and structural data processing
\end{itemize}

\hypertarget{azure-to-sensors-a2s-connectivity-deployment-field-level-connectivity}{%
\section{Azure to Sensors (A2S) connectivity deployment (field level
connectivity)}\label{azure-to-sensors-a2s-connectivity-deployment-field-level-connectivity}}

\begin{quote}
The title must be revised
\end{quote}

\begin{itemize}
\tightlist
\item
  \textbf{Architecture} - Domain model presenting relationship between
  the: Azure, PubSub Gateway, Device, Design and development tools
\item
  \textbf{Connectivity} - Describe reactive nature of the Azure
  monitoring process data (telemetry) services.
\item
  \textbf{Deployment phases}

  \begin{itemize}
  \tightlist
  \item
    Design
  \item
    Gateway and devices registration
  \item
    Authentication
  \item
    Device/Service association
  \item
    Device/Application association
  \item
    Establishing session

    \begin{itemize}
    \tightlist
    \item
      Device/Device Template (Device Capability Model) association -
      establishing a semantic-context
    \item
      Security management - establishing security-context
    \end{itemize}
  \item
    Interconnection - exchange of data
  \item
    Maintenance
  \end{itemize}
\end{itemize}

We have selected
\href{https://docs.microsoft.com/azure/iot-central/core/}{IoT Central}
because:

\begin{itemize}
\tightlist
\item
  provides process data visualization user interface
\item
  allows to describe devices using metadata containing telemetry data
  types
\end{itemize}

\hypertarget{gateway-implementation}{%
\section{Gateway implementation}\label{gateway-implementation}}

\begin{itemize}
\tightlist
\item
  Architecture
\item
  Protocol selection and mapping
\item
  Configuration
\item
  Testing
\end{itemize}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The OPC UA PubSub to Azure gateway (\texttt{AzureGateway})
implementation has been just published on GitHub as the open-source (MIT
licensed) as a part of the more general concept of the Object-Oriented
Internet reactive networking. It is proof of the concept that

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  OPC UA PubSub can be implemented as a powerful standalone package - no
  C/S dependency
\item
  Azure interoperability can be implemented as an out-of-band
  communication (MQTT, AMQP, HTTP) - no PubSub dependency
\item
  Process data functionality can be composable at run-time - no
  programming required
\end{enumerate}

\end{document}