### 2020-09-29 Latex/Manuscript..tex
Workout Markdown\10.DocumentHeader.md #25
fixes #25

\documentclass{article}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{stfloats}
\usepackage{url}
\usepackage{authblk}

\providecommand{\tightlist} { \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\providecommand{\keywords}[1]{  \small	  \textbf{\textit{Keywords---}} #1 }

\title{Object-Oriented Internet - Azure interoperability}
% \title{Object-Oriented Internet - reactive visualization of asynchronous data using AZURE}
\author[1]{Mariusz Postół}
\author[2]{Piotr Szymczak}
\author[3]{Clemens Vasters}
\affil[1, 2]{Institute of Information Technology, Lodz University of Technology, Łódź, Poland}
% \email{mailto:mariusz.postol@p.lodz.pl} 
\affil[3]{Microsort}
%\tocauthor{Mariusz Postół}
%\institute{ Institute of Information Technology, Lodz University of Technology, Łódź, Poland \email{mailto:mariusz.postol@p.lodz.pl} }

\begin{document}

\maketitle

\begin{abstract}

TBD

\keywords{Azure, Cloud Computing, Object-Oriented Internet, Industrial communication, Industry 4.0, Internet of Things, Machine to Machine Communication, OPC Unified Architecture}

\end{abstract}

\maketitle

\hypertarget{introduction}{%
  \section{Introduction}\label{introduction}}

\begin{itemize}
  \tightlist
  \item
        \textbf{Subject} - A basic matter of thought, discussion,
        investigation, development, etc. Describe the problem and the
        motivation for undertaking the effort to solve the problem.
  \item
        \textbf{Goal} What we are going to achieve - the result or achievement
        toward which effort is directed.
  \item
        \textbf{Scope} - What we must do to prove the goal have been achieved.
        Extent or range of development, view, outlook, application, operation,
        effectiveness, etc.
  \item
        \textbf{Related work} - Any information about available reusable
        deliverables related to this work.
\end{itemize}

\hypertarget{azure-main-technology-features}{%
  \section{Azure Main Technology
    Features}\label{azure-main-technology-features}}

\begin{itemize}
  \tightlist
  \item
        \textbf{Selection of the service}
  \item
        \textbf{Metadata} - must be discussed in context of the design/run
        time stages.

        \begin{itemize}
          \tightlist
          \item
                \textbf{Device Template (DT)}
          \item
                \textbf{Device Capability Model}
          \item
                \textbf{Interface}
          \item
                \textbf{Digital Twin Definition Language (DTDL)}
        \end{itemize}
  \item
        \textbf{Simple, complex and structural data processing}
  \item
        \textbf{Connectivity}
  \item
        \textbf{How to implement} All about available libraries and tools
\end{itemize}

\hypertarget{ooi-main-technology-features}{%
  \section{OOI Main Technology
    Features}\label{ooi-main-technology-features}}

\begin{itemize}
  \tightlist
  \item
        Machinie To Machine communication based on the semantic data
  \item
        OOI PubSub Implementation Architecture
  \item
        Simple, complex and structural data processing
\end{itemize}

\hypertarget{azure-to-sensors-a2s-connectivity-deployment-field-level-connectivity}{%
  \section{Azure to Sensors (A2S) connectivity deployment (field level
    connectivity)}\label{azure-to-sensors-a2s-connectivity-deployment-field-level-connectivity}}

\begin{quote}
  The title must be revised
\end{quote}

\begin{itemize}
  \tightlist
  \item
        \textbf{Architecture} - Domain model presenting relationship between
        the: Azure, PubSub Gateway, Device, Design and development tools
  \item
        \textbf{Connectivity} - Describe reactive nature of the Azure
        monitoring process data (telemetry) services.
  \item
        \textbf{Deployment phases}

        \begin{itemize}
          \tightlist
          \item
                Design
          \item
                Gateway and devices registration
          \item
                Authentication
          \item
                Device/Service association
          \item
                Device/Application association
          \item
                Establishing session

                \begin{itemize}
                  \tightlist
                  \item
                        Device/Device Template (Device Capability Model) association -
                        establishing a semantic-context
                  \item
                        Security management - establishing security-context
                \end{itemize}
          \item
                Interconnection - exchange of data
          \item
                Maintenance
        \end{itemize}
\end{itemize}

We have selected
\href{https://docs.microsoft.com/azure/iot-central/core/}{IoT Central}
because:

\begin{itemize}
  \tightlist
  \item
        provides process data visualization user interface
  \item
        allows to describe devices using metadata containing telemetry data
        types
\end{itemize}

\hypertarget{gateway-implementation}{%
  \section{Gateway implementation}\label{gateway-implementation}}

\begin{itemize}
  \tightlist
  \item
        Architecture
  \item
        Protocol selection and mapping
  \item
        Configuration
  \item
        Testing
\end{itemize}

\hypertarget{conclusion}{%
  \section{Conclusion}\label{conclusion}}

The OPC UA PubSub to Azure gateway (\texttt{AzureGateway})
implementation has been just published on GitHub as the open-source (MIT
licensed) as a part of the more general concept of the Object-Oriented
Internet reactive networking. It is proof of the concept that

\begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
        OPC UA PubSub can be implemented as a powerful standalone package - no
        C/S dependency
  \item
        Azure interoperability can be implemented as an out-of-band
        communication (MQTT, AMQP, HTTP) - no PubSub dependency
  \item
        Process data functionality can be composable at run-time - no
        programming required
\end{enumerate}

\bibliography{Manuscript}
\bibliographystyle{splncs04}

\end{document}